\subsection{Background: EMF, GMF and Epsilon}
\label{sec:eclipse-based-dsls}
Many people consider the core of a language to be its abstract
syntax. The Eclipse Modelling Framework (EMF)~\cite{steinberg08} is a
modelling framework and code generation facility for building tools
and other applications based on a structured data model. Part of this
framework includes Ecore~\cite{steinberg08} which is a UML class
diagram like language for describing meta models for DSLs. Such a
model serves as the basis for creating a graphical syntax for a DSL
using the graphical modelling framework. The Graphical Modelling
Framework (GMF)~\cite{gronback09} provides features allowing one to
develop, from an Ecore meta model, a graphical concrete syntax for a
DSL. This editor can be used to produce model instances of the DSL
described by the underlying Ecore meta model. Often, the development
of a GMF editor is motivated by the possibility of producing, from an
instance model created by the editor, some sort of output usually in
the form of text or program code. To help with this task, users can
make use of what are known as model transformations. For this task, we
make use of Epsilon~\cite{kolovos2012}. Epsilon provides a family of
languages and features for defining and applying model
transformations, comparisons, validation and code generation.


%  We use the following: the Epsilon
% object language (EOL) that provides a common set of model management
% constructs. The Epsilon transformation language (ETL) for specifying
% model to model transformations; The Epsilon generation language (EGL)
% for model to text transformations; and the Epsilon wizard
% language(EWL) for defining and executing transformation workflows,
% including activating transformations from a GMF editor.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
